Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic

### 2022-04-05 psy_peds_3_umich.Rmd
added the rmd run by umich

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-06-05"
time_period <-"months" #possible values: days, weeks, months
```

@@

getwd()
files <- readInputFiles( path      = folder_4ce_files, 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )

@@

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV3.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   dplyr::select( disorder_group, ICD10_Code, description )
```

@@

### Calculating age at hospitalization time
```{r}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
  inner_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
            by = c("patient_num", "days_since_admission"))
obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
```

@@

bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list,
                                          .id = "iteration")

z <-  mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]) / sd(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])

pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)

plot_annotation <- paste0("Coeffs differences: ",
       round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),
       " (pvalue: ",
       pvalue_coeff_interaction,
       ")")

@@

```{r}
# counts per time period and sex
patients_hospitalized_agg_sex <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
           sex %in% c("female", "male")) %>%
    group_by( time_p, sex ) %>%
  summarise(count_sex =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))

>>>>

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/Users/hanauer/Desktop/DONT_INDEX/4ceData/Input/phase22_11NOV2021/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
data_update_date <- "2021-11-04"
time_period <-"months" #possible values: days, weeks, months
```

@@

files <- readInputFiles( path      = folder_4ce_files, 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )

@@

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   dplyr::select( disorder_group, ICD10_Code, description )
```

@@

### Calculating age at hospitalization time
```{r}
# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
            by = c("patient_num", "days_since_admission"))
stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
obs_raw_age <- obs_raw_age[ !is.na( obs_raw_age$calendar_date), ]
obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
```

@@

bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
mean(bootstrapped_coefficients_df$time)
mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
pvalue_coeff_interaction <- ifelse(pvalue_coeff_interaction == 0,
                                   "< 1e-4",
                                   str(pvalue_coeff_interaction))
plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",
                     round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),
                     " (pvalue: ",
                     pvalue_coeff_interaction,
                     ")")

@@

```{r}
# counts per time period and sex
all_codes_psy$sex <- trimws( all_codes_psy$sex)
all_codes_psy$race_4ce <- trimws( all_codes_psy$race_4ce)
patients_hospitalized_agg_sex <- all_codes_psy %>%
  filter(date < end_date_plots &
         date >= start_date_plots & 
           sex %in% c("female", "male")) %>%
    group_by( time_p, sex ) %>%
  summarise(count_sex =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
                               n_distinct(patient_num), 
                               0.5))
